#R script file for tutorial on a forecast accuracy study.
#The data are in an SQLite database file called "NewExtract". The product information has been disguised.


# Measuring Forecast accuracy

### Weighted Average Percent Error (WAPE) 
$$
100r = \frac{\sum_{i=1}^N |\hat{x}_i - x_i|}{\sum_{i=1}^N x_i}
$$

### Moving average
$$
\hat{x}_{x+1} = \displaystyle \Sigma_{i=0}^{s-1} w_i x_{n-i}
$$

### Exponential Smoothing
The weights need to add up to one and can be defined as unequal. We should consider weighting more recent months more strongly. Each exponential smoothing is defined with an $\alpha$. You can use the previous value in the computation.

$$
\begin{align}
\hat{x}_{n+1}
&= \alpha x_n + (1-\alpha) \hat{x}_n \\
&= \alpha \sum_{i=0}^{n-2} (1-\alpha)^i x_{n-i} 
                        + (1-\alpha)^{n-1} x_1
\end{align}
$$

The exponential weights do sum to one.

### Double Exponential Smoothing
Each double exponential smoothing is defined with an $\alpha$ and $\beta$.
Double exponential smoothing captures the changing trend.

**Please know how to fill up the table.**



Given the value of alpha, you minimise the SSE.

####################### Part I: Link R to SQL Starts ######################################
```{r}
#### Prep: Load R package; Set working directory; Load the database to R
# Remove all variables from the R environment to create a fresh start
rm(list=ls())

# We next load some packages that are generally useful for data analytics.
#In each case, check first to see if a required package is installed. If not, then install it 
# and invoke it as a library.
# if(!require(RSQLite)){
#   install.packages("RSQLite")
#   library(RSQLite)
# }
# if(!require(treemap)){
#   install.packages("treemap")
#   library(treemap)
# }
# if(!require(stats)){
#   install.packages("stats")
#   library(stats)
# }
AirPassengers
plot(AirPassengers)
plot(HoltWinters(AirPassengers))
plot(HoltWinters(AirPassengers, beta=FALSE))
plot(HoltWinters(AirPassengers, gamma=FALSE))
plot(HoltWinters(AirPassengers, gamma=FALSE, beta=FALSE))
```

# Loading SQL database

```{r}
# Similarly, we need to change the working directory to the directory where we want to read and save data
# Define a variable, wd, with a string describing the path to the directory where we want to read and save data
# Change this to meet your needs.!!!!!!!!!

# Load the data base to R and test the RSQLite library. Put them inside a function so they are executed  only
# if the function is called and ignored otherwise.

testRSQLite <- function(){
  # Capture the name of the database file (we assume it is located in the working directory)
  dbname <- "wk5b-time_series.sqlite"
  # Establish a database connection to this database file
  conn <- dbConnect(SQLite(),dbname)
  # Take a look at the structure of this connection: it should display the defaults, such as loadable.extensions=TRUE
  str(conn)
  # Assuming there is no error in connecting, we try now to get some data from the database.
  # Define an SQL query to get a small amount of data (use SQLite syntax)
  query <- "SELECT * FROM NewExtract LIMIT 20"
  # Use the connection to get the query results
  queryresult <- dbGetQuery(conn,query)
  # Close the connection to the database
  dbDisconnect(conn)
  # Return the result (it is in the form of a data.frame)
  queryresult
}
# Run the next statement (without the comment #) if you want a quick test
testRSQLite()
```


```{r}
#### Execute SQL queries in R
# Let's create a function to grab data from our database according to a query we provide. This will be useful.
getExtract <- function(query){
  # Capture the name of the database file (we assume it is located in the working directory)
  dbname <- "wk5b-time_series.sqlite"
  # Establish a database connection to this database file
  conn <- dbConnect(SQLite(),dbname)
  # Use the connection to get the query results
  queryresult <- dbGetQuery(conn,query)
  # Close the connection to the database
  dbDisconnect(conn)
  # Return the result (it is in the form of a data.frame)
  queryresult
}
# Run the next statement to test the getExtract function
getExtract("SELECT * FROM NewExtract LIMIT 20")
```


```{r}
#Execute Statements One at a Time
# Define a query to aggregate actual sales by region
query <- "select region,sum(actual) as TotalActual from NewExtract group by region order by sum(actual) desc"
# Grab the data into a data.frame
result <- getExtract(query)
# Display the result
result

# Plot the values
plot(result$TotalActual)
# Plot the values as a bar chart
barplot(result$TotalActual)
# Plot the values as a bar chart with labels
barplot(result$TotalActual,names.arg=result$Region)
# Plot the values as a treemap with labels
treemap(result,title="Total Quantity by Region",index="Region",vSize="TotalActual")
# Which plot pattern do you like best?

####################### Part I: Link R to SQL Starts Ends######################################
```


```{r}
####################### Part II: Forecasting in R Starts ######################################
# In part II, we will run different kinds of forecast for a specific product in the database

# First, we grab the data for the specific product

####################### Grab data for a specific product Starts ######################################
# Let's create a function to grab the data for a specific product from our database
getExtractProduct <- function(region, segment, family, subfamily, beforepkg, afterpkg){

  # Let's form a query to get all of the data for this product
  query <- paste ("Select * from NewExtract where region ='",
                  region,"' and Segment ='",
                  segment, "' and Family ='",
                  family,"' and SubFamily='", 
                  subfamily,"' and BeforePkg='",
                  beforepkg,"' and AfterPkg='",
                  afterpkg,"'",
                  sep="")

  # Let's tack on an 'order by' clause to make sure the sort is correct
  query <- paste(query," order by year,monthseq", sep="")

  # Grab the data into a data.frame
  result <- getExtract(query)

  # return the result
  result
}
```


```{r}
# call the function getExtractProduct to grab the data for a specific product

# Let's focus on region R031 (Why?)
region <- "R031"

# Let's focus on segment S17, family F033 and subfamily SF101
segment <- "S17"
family <- "F033"
subfamily <- "SF101"

# Let's focus on BeforePkg B345 and AfterPkg A001 as our final choice
beforepkg <- "B345"
afterpkg <- "A001"

# Grab all the data for this product using function getExtractProduct defined above
result <- getExtractProduct(region, segment, family, subfamily, beforepkg, afterpkg)
result
```

```{r}
# Note that the database has provided the company's forecast in the column "forecast"
# We first plot the actual demand against the company's forecast for comparison
forecasts <- result$forecast
actuals <- result$actual
plot(actuals)
# Connect the points with lines so the sequence is clear
lines(actuals)
# Now add the corporate forecasts as red circles
points(forecasts,col=c("red"))
# How good are the forecasts? Perfection in the next plot would be a straight line angled at 45 degrees
# You could also plot the actual against forecast by uncomment the command below
# plot(actuals,forecasts)
# So, the forecasts are not good. 
```



```{r}
# Let's develop a quantitative measure of forecast accuracy.
errors <- forecasts-actuals
# Take the absolute difference of forecasts versus actuals.
abserrors <- abs(errors)
# Compute the total of the absolute errors
totalerror <- sum(abserrors)
# Compute the total actual demand
totalactual <- sum(actuals)
# The ratio is the relative absolute error of forecast
if (totalactual>0) {
    relativeabsoluteerror <- totalerror/totalactual
} else {
  relativeabsoluteerror <- 0
}
# Practitioners multiply the relative absolute error this by 100 to get what they call the WAPE (Weighted Average Percent Error)
WAPE <- relativeabsoluteerror*100
WAPE
# So, the Weighted Average Percent Error is about 49%. That is pretty high. 
# And this is the largest selling product in the largest segment in the largest region!
# The forecast errors for smaller selling products will likely be much worse.
####################### Grab data for a specific product Ends ###################################### 
```

```{r}
# Next we try different forecasting methods in R
############################ Moving Average Forecasting Starts  ###################################################
# The moving average technique takes a vector of weights over a certain number of periods. Let's do a moving average over three periods.
weights <- rep(1/3,3)
weights
# There are many ways to implement moving average in R. The filter method is easy. 
# The argument "Sides=1" says to just use data to the left of point to be forecast.
maforecasts <- round(filter(actuals,weights,method="convolution",sides=1,circular=FALSE))
# The first few fitted values are NA because there is insufficient history to calculate the moving average.
# Omit NA entries.
maforecasts <- na.omit(maforecasts)
# The resulting structure is a time series. We just want the fitted values
class(maforecasts)
forecasts <- as.numeric(maforecasts)
str(forecasts)
```

```{r}
# Plot the data
# Omit the first few entries in actuals which has not predictions
actualsnew <- tail(actuals, length(forecasts))
plot(actualsnew)
# Connect the points with lines so the sequence is clear
lines(actualsnew)
# Now add the corporate forecasts as red circles
points(forecasts,col=c("red"))
title("Moving Average 3")
```

```{r}
# Calculate the quantitative measure of forecast accuracy.
errors <- forecasts-actualsnew
# Take the absolute difference of forecasts versus actuals.
abserrors <- abs(errors)
# Compute the total of the absolute errors
totalerror <- sum(abserrors)
# Compute the total actual demand
totalactual <- sum(actualsnew)
# The ratio is the relative absolute error of forecast
if (totalactual>0) {
    relativeabsoluteerror <- totalerror/totalactual
  } else {
  relativeabsoluteerror <- 0
}
# Practitioners multiply the relative absolute error this by 100 to get what they call the WAPE (Weighted Average Percent Error)
WAPE <- relativeabsoluteerror*100
WAPE
# So, the Weighted Average Percent Error is about 22%. That is much better.
############################  Moving Average Forecasting Ends ###################################################
```

```{r}
# We try another powerful technique Holt-Winters method.
############################  Holt-Winters method Starts ###################################################

################# Get read for Holt_Winters #################
# Holt-Winters requires a time series structure. So we convert the actuals vector into a time series by specifying the frequency (in this case, the number of months per year)
tsactuals <- ts(actuals,frequency=12)
str(tsactuals)
plot(tsactuals)
```

```{r}
# Here we try three Holt-Winters methods
# We first try Holt-Winters with just exponential smoothing (beta=False,gamma=FALSE)
################# Holt-Winters with just exponential smoothing (beta=False,gamma=FALSE) #################
hw <- HoltWinters(tsactuals, beta=FALSE,gamma=FALSE)
# THe Holt Winters data structure is complicated. We want to extract the first column of the fitted values.
hwfitted <- fitted(hw)[,1]
# The fitted values are a time series so we convert it to numeric (and round them to whole numbers for simplicity).
forecasts <- round(as.numeric(fitted(hw)[,1]))
```

```{r}
# Here we try three Holt-Winters methods
# We first try Holt-Winters with just exponential smoothing (beta=False,gamma=FALSE)
################# Holt-Winters with just exponential smoothing (beta=False,gamma=FALSE) #################
hw <- HoltWinters(tsactuals, beta=FALSE,gamma=FALSE)
# THe Holt Winters data structure is complicated. We want to extract the first column of the fitted values.
hwfitted <- fitted(hw)[,1]
# The fitted values are a time series so we convert it to numeric (and round them to whole numbers for simplicity).
forecasts <- round(as.numeric(fitted(hw)[,1]))
```



```{r}
# Here we try three Holt-Winters methods
# We first try Holt-Winters with just exponential smoothing (beta=False,gamma=FALSE)
################# Holt-Winters with just exponential smoothing (beta=False,gamma=FALSE) #################
hw <- HoltWinters(tsactuals, beta=FALSE,gamma=FALSE)
# THe Holt Winters data structure is complicated. We want to extract the first column of the fitted values.
hwfitted <- fitted(hw)[,1]
# The fitted values are a time series so we convert it to numeric (and round them to whole numbers for simplicity).
forecasts <- round(as.numeric(fitted(hw)[,1]))
```

```{r}
# Plot the Holt-Winters forecasts
# The fitted values are missing the first entry, so we the first entries in actuals which has not predictions
actualsnew <- tail(actuals, length(forecasts))
plot(actualsnew)
# Connect the points with lines so the sequence is clear
lines(actualsnew)
# Now add the corporate forecasts as red circles
points(forecasts,col=c("red"))
title("Exponential Smoothing")
```

```{r}
# Calculate the quantitative measure of forecast accuracy.
errors <- forecasts-actualsnew
# Take the absolute difference of forecasts versus actuals.
abserrors <- abs(errors)
# Compute the total of the absolute errors
totalerror <- sum(abserrors)
# Compute the total actual demand
totalactual <- sum(actualsnew)
# The ratio is the relative absolute error of forecast
if (totalactual>0) {
    relativeabsoluteerror <- totalerror/totalactual
} else {
  relativeabsoluteerror <- 0
}
# Practitioners multiply the relative absolute error this by 100 to get what they call the WAPE (Weighted Average Percent Error)
WAPE <- relativeabsoluteerror*100
WAPE
```


```{r}
# We secondly try Holt-Winters without seasonality (double exponential smoothing (gamma=FALSE)
################# Holt-Winters without seasonality (gamma=FALSE) #################
# Now call the Holt-Winters function without seasonality (gamma=FALSE). This is equivalent to double exponential smoothing.
hw <- HoltWinters(tsactuals, gamma=FALSE)
# THe Holt Winters data structure is complicated. We want to extract the first column of the fitted values.
hwfitted <- fitted(hw)[,1]
# The fitted values are a time series so we convert it to numeric (and round them to whole numbers for simplicity).
forecasts <- round(as.numeric(fitted(hw)[,1]))

####################Input your code for plotting and calculating WAPE ##########
```

```{r}
# Finally we try the full Holt-Winters
################# Holt-Winters Forecasting #################
####################Input your code for forecasting using Holt-Winters ##############
 

 

####################Input your code for plotting and calculating WAPE ##########





###############################################################################
# Out of all the four forecasting methods above, which one gives the lowest WAPE? 
############################  Holt-Winters method Ends ###################################################

####################### Part II: Forecasting in R Ends ######################################
```

