---
title: "Week 1 Practice"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
--- 

Load the dataframe mtcars available with the base R installation in data. The data was extracted from the 1974 Motor Trend US magazine, and comprises 
- fuel consumption 
- 10 aspects of automobile design
- performance 
for 32 automobiles (1973-74 models).


```{r}
library(ggplot2)
head(mtcars)
```

(a) Use ggplot() to plot the weight versus miles per gallon and comment on the relationship.
Add to this plot, a coloring of the points based on the number of cylinders, and the sizing of the points based on the displacement (volume) of the car.


```{r}
acol <- ggplot(mtcars,aes(x=wt, y=mpg, color=cyl, cex=disp))
acol + geom_point(na.rm=T) + geom_smooth(na.rm=T)
```

(b) Use the tapply() function to compute the standard deviation of the mpg for groups with the same number of cylinders

Refer to: https://www.rdocumentation.org/packages/base/versions/3.6.1/topics/tapply

```{r}
tapply(mtcars$mpg, mtcars$cyl, sd)
```

