---
title: "WHO Notebook"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
--- 
```{r}
# install.packages("ggplot2") 
```

### Comments
Compared to Python matplotlib, you add characteristics by adding `+`.

### Guide

Please add the file into the folder.

One of the advantages of software such as R is that you can add and install packages that provided added functionality. The `ggplot2` packages is one such package that improves on the basic plotting in R. In this package, you build a graph by adding layer upon layer where you specify the data, aesthetic mapping, statistical transformations (if any) and geometric object (type of plot).

We will use a data set from <http://www.who.int/gho/en>. The dataset contains data on 194 countries from WHO where the variables are: 
1. Country - name of country, 
2. Region - region the country lies in, 
3. Population - population in thousands, 
4. Under15 - Percentage of the population under 15 years of age, 
5. Over60 - Percentage of the population over 60 years of age,
6. FertilityRate - Average number of children per woman,
7. LifeExpectancy - Life expectancy in years,
8. LiteracyRate - Literacy rate among adults at least 15 years of age,
9. GNI - Gross national income per capita


First we have a look at the data set
```{r}
library(ggplot2)
WHO <- read.csv("w1c2b-WHO.csv")
```

```{r}
str(WHO)
```

`which` to find the index.

```{r}
index <- which(WHO$Country == "Singapore")
index
WHO[index,]$GNI
summary(WHO$Over60)
summary(WHO$Under15)
```


Let us see how many countries are there in each region
```{r}
bar <- ggplot(WHO) + 
  geom_bar(mapping = aes(x = Region, fill = Region), 
           show.legend = FALSE, width = 1) + 
  theme(aspect.ratio = 1) + 
  labs(x = NULL, y = NULL)
bar + coord_flip()
bar + coord_polar()
```


### Scatter plots

Now we compare the following

1. GNI vs Fertility rate
2. GNI vc Literacy Rate
3. GNI vc Life Expectancy

We also subgroup by region sometimes

`facet_wrap` to organise the chart more neatly. Probably should display region with marker color (later), and population size with marker size.

```{r}
a <- ggplot(WHO, aes(x=GNI, y=FertilityRate))
a + geom_point(na.rm=T) + facet_wrap(.~Region)
```

```{r}
a + geom_point(na.rm=T) + geom_smooth(na.rm=T)
```


Plots a confidence boundary (?) for countries of every region. Messy but we just do it. 

(Is there some statistical way to constrain that it is impossible to achieve negative fertility rate?)

```{r}
acol <- ggplot(WHO,aes(x=GNI, y=FertilityRate, color=Region))
acol + geom_point(na.rm=T)
acol + geom_point(na.rm=T) + geom_smooth(na.rm=T)
```



