---
title: 'Basic R Notebook: Lecture 2'
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

### Comparision on R and Jupyter
Here you are editing the markdown script of the notebook. This is not very possible on Jupyter, which is saved as a JSON.
On RStudio you can also collaspe the section, makeing your view more focused.


### Notes on MacOS
Keyboard shortcuts
https://bookdown.org/yihui/rmarkdown/notebook.html

Cmd + Opt + I for new cell
Refer to the the 'Run' option on the top right of this window to see the shortcut related to running and restarting the cell.
- Restart and run all
- Run chunk
- Run line



### Other notes
?(command) for help

### R class types

R stores objects in a variety of classes - numeric, integer, character, logical, list, matrix, dataframe and has logical overriding operations when you convert from one class to another.

"integer" class is part of "numeric"

```{r}
z <- c(0:9)
z

class(z)
```


```{r}
z1 <- c("a","b","d")
z1

class(z1)

```

The result returns as NA if the data cannot be casted.

```{r}
w <- as.character(z1)
w

class(w)
as.integer(w)
```

### Factors
 
Essentially tabulate by count?

```{r}
ct <- c("jap", "kor", "kor", "jap", "sgp")
class(ct)
fct <- factor(ct)
fct
summary(fct)
```


The function tapply:

# Matrix and arrays

### Matrix operations

```{r}
r <- matrix(c(3,4,5,6,7,8), nrow=3, ncol=2)
r

class(r)
dim(r)
r[2,2]
r[5]
```

r[5] reads the flatted array. (Different from Python)

### Array operations

Casting an array into your specified multi-dimensional matrix.

```{r}
s <- array(c(3,4,5,6,7,8), c(3,2))
s

class(s)
z <- 1:50
dim(z) <- c(5,2,5)
```

### Other operations with matrices

Writing an identity matrix of size 10

```{r}
diag(10)
```


### Matrix multiplication

`%*%` is matrix multiplication. Probably there are some commands for transpose, etc.

```{r}
a <- array(c(2,1,-1,2), c(2,2))
a%*%a
```


### Solutions of linear equations

solve(a) will invert the matrix. Please revise your linear algebra.

```{r}
a <- array(c(2,1,-1,2), c(2,2))
b <- c(4,4)
a
b
a%*%solve(a)
```


## Eigen decomposition of a matrix

Eigen decomposition which values and vectors. Use the dolar sign to get its characterisitics.

```{r}
eigen(a)
eigen(a)$vectors
```

## List (Ordered Collection)

List consists of an ordered collection of objects that can be of different or the same type.

We are defining a list/dictionary of "objects". 

```{r}
barack <- list(age=58, gender="M", child.ages=c(21,18))
barack
class(barack)
barack[1]
class(barack[1])
barack[[1]]
class(barack[[1]])
```

```{r}
serena <- list(age=38, gender="F", child.ages=2)
serena
```

Here we are combining the arrays together? This looks so unwieldy.

```{r}
Cleg <- c(barack, serena)
Cleg
Cleg[1]
Cleg[4]
```

## Dataframes

Dataframes are a tightly coupled collection of variables that share many of the properties of matrices and lists and is the fundamental data structure that will be used in most of this course.

```{r}
CELG <- data.frame(names=c("barack","serena"), ages=c(58,38),children=c(2,1))
CELG
```

```{r}
CELG$spouse <- c("michelle","alexis")
CELG
```

