######## Data Preparation
```{r}
# Load the data on the supreme court
supreme <- read.csv("wk8-law.csv") # We have 623 observations and 20 variables
head(supreme) # First part of the dataframe
# str(supreme) # Internal structure of the dataframe 
# summary(supreme) # Summary of the data

# A few comments on the data:
# docket: case number
# term: year the case was discussed
# party_1 and party_2: parties involved in the case
# rehndir stevdir ocondir scaldir kendir soutdir thomdir gindir brydir: direction of the judgement of the 9 judges
# (i.e., Rehquist, Stevens, O'Connor, Scalia, Kennedy, Souter, Thomor, Ginsburg, Breyen). 0 means liberal, 1 conservative. 
# 9 indicates that the vote is not available for the case.
# petit: petitioner type
# respon: respondant type
# circuit: circuit of origin
# uncost: binary number indicating if the petitioner argued constitutionality of a law or practice
# lctdir: lower court direction of results (liberal or conservative)
# issue: area of issue
# result: 0 means liberal, 1 conservative.

# Let's now focus on a specific judge, say Stevens (we remove data with no entry from Stevens--Stevens was present)
stevens <- subset(supreme[,c("docket","term","stevdir","petit","respon","circuit","unconst","lctdir","issue","result")],supreme$stevdir!=9)

# Processing the output result to affirm or reverse for judge Stevens
stevens$rev <- as.integer((stevens$lctdir=="conser" & stevens$stevdir==0) | (stevens$lctdir=="liberal" & stevens$stevdir==1))
```


```{r}
# This creates a new variable ($rev) in the dataframe that takes a value of 1 if Stevens decision reverses the lower court decision
# and 0 if the judge affirms the lower court decision. Note that a similar analysis can be done for the other judges or for the overall case result.
table(stevens$rev)
```

######## Modelling exercise: logistic regression
```{r}
# Let's prepare the data for our modelling exercise
# Load the caTools package (basic utility functions)
if(!require(caTools)){
  install.packages("caTools")
  library(caTools)
}
# Set the seed
set.seed(1)
# Create train and test sets (with balanced response from the judge Stevens)
spl <- sample.split(stevens$rev,SplitRatio=0.7) # We use 70% of the data for training
train <- subset(stevens,spl==TRUE); # table(train$rev)
test <- subset(stevens,spl==FALSE); # table(test$rev)
# There is only one realization of the IR value (Interstate Relations), which does not
# appear in the training dataset. Let's thus modify the test dataset as follows:
test <- subset(test,test$issue!="IR")

# Logistic regression
# We predict judge Stevens' decision 
m1 <- glm(rev~petit+respon+circuit+unconst+issue+lctdir,data=train,family="binomial")
summary(m1)
# let's now try to make a prediction on the test data
p1 <- predict(m1,newdata=test,type="response")

# How does the model perform?
table(p1>=0.5,test$rev)
# Confusion matrix
#        0  1
# FALSE 53 32
# TRUE  29 70
# The accuracy is about 0.66 

# Area Under the Curve
if(!require(ROCR)){
  install.packages("ROCR")
  library(ROCR)
}
pred <- prediction(p1,test$rev) # create the prediction object
perf <- performance(pred, x.measure="fpr",measure="tpr") # calculate False and True Positive Rate (FPR and TPR)
plot(perf) # FPR vs TPR
performance(pred,measure="auc") # 0.7404352


######## Classification And Regression Trees (CARTs)
```

```{r}
# rpart is for CART
if(!require(rpart)){
  install.packages("rpart")
  library(rpart)
}
# rpart.plot extend some functionalities of rpart
if(!require(rpart.plot)){
  install.packages("rpart.plot")
  library(rpart.plot)
}
# rattle gives additional functionalities
if(!require(rattle)){
  install.packages("rattle")
  library(rattle)
}
# colors
if(!require(RColorBrewer)){
  install.packages("RColorBrewer")
  library(RColorBrewer)
}
```

```{r}
# Learning the tree (when building the tree for classification, rpart uses the Gini index)
#
# Correct command
cart1 <- rpart(rev~petit+respon+circuit+lctdir+issue+unconst,
               data=train, 
               method="class")
```

```{r}
# More details on the tree structure
#
# With this command, we get information on: split criteria, no. rows (observations) in this node, no. of misclassified, predicted class,
# and % of rows in predicted class for this node.
# node), split, n, loss, yval, (yprob)
#      * denotes terminal node
print(cart1)
```

```{r}
# Tree visualization
# 
# Let's first use the default function implemented in the rpart package
plot(cart1); text(cart1)
#
# Let's use a more advanced function from the rpart.plot package
```

```{r}
prp(cart1) 
prp(cart1,type=1) # labels all nodes (not just the leaves)
```

```{r}
# There many other visualization options: see an example with the fancyRpartPlot() function, which requires rpart.plot, rattle, and RColorBrewer.
fancyRpartPlot(cart1)
```

```{r}
# Predictions
#
# The function predict uses cart1 and the data test to carry out the predictions.
# The function chooses the class with the highest probability.
predictcart1 <- predict(cart1, newdata = test)
#
# Confusion matrix
# with a certain defined cut-off
#...

# The accuracy is ...
# We can see that this model has performance similar to that achieved by the logistic regression
#

# We can also use the Area Under the Curve
predictcart1_prob <- predictcart1
pred_cart1 <- prediction(predictcart1[,2], test$rev)
perf_cart1 <- performance(pred_cart1,x.measure="fpr",measure="tpr")
plot(perf_cart1)
as.numeric(performance(pred_cart1,measure="auc")@y.values)
```


######## Pruning
```{r}
# Display the cp table for model cart1
printcp(cart1)

# We can also plot the relation between cp and error
plotcp(cart1)
```

```{r}
# We are now ready for pruning

# Pruning. Which value of pruning would you use?
cart2 <- prune(cart1,cp=0.01)
fancyRpartPlot(cart2)
predictcart2 <- predict(cart2,newdata=test,type="class")
table(test$rev,predictcart2)
```


######## Random Forests

**Intuition of ensemble training** - each person have different knowledge, and different logical framework. Individual trees may not be accurate (and are faster to train), but make the ensembled prediction is more accurate. (Analogy - Who Wants to be a Millionaire)

Sample the training dataset with replacement (boosting or bagging)
- Build a classification or regression tree
- When splitting a node
- Randomly choose $m$ parameters
- Find the best cut-point by exhasutive search
- Stop when the exit condition is met

Training a tree in a forest is faster than training a CART because only a subset of predictors are used.



**Meta-parameters**

The training does not optimise the meta-parameters.

- $T^*$ number of tree in the ensemble `ntree` (recommended 100 to 1000)
- $m$ = number of predictors chosen when creating a split `mtry` (recommended sqrt(p) for classification, p/r for regression)
- $n_{min}$ minimum number of nodes in each leaf `nodesize` (recommended 2 to 20)



Downsides (compart to CART)

- slower (though parallelizable across cores)
- less interpretable





```{r}
# Load the randomForest package (it can be used for both classification and regression)
if(!require(randomForest)){
  install.packages("randomForest")
  library(randomForest)
}
# set seed
set.seed(1)
#
# Build a forest of 200 trees, with leaves 5 observations in the terminal nodes
forest <- randomForest(as.factor(rev)~petit+respon+circuit+unconst+lctdir+issue,
                       data=train, nodesize=5, ntree=200)
forest
# The prediction is carried out through majority vote (across all trees)
predictforest <- predict(forest, newdata = test, type="class")
table(predictforest, test$rev)

# Variable importance
varImpPlot(forest)
# forest$importance

```